---
title: "Assignment-2"
author: "TarakRam Nunna"
date: "10/11/2021"
output:
  word_document: default
  html_document: default
---

# Setting working directory
```{r}
getwd()
setwd("C:/Users/TARAKRAM/OneDrive/Desktop/Business Analytics/Assignment-2")
```

### 1) Run the following code in R-studio to create two variables X and Y. 
```{r}
set.seed(2017) 
X=runif(100)*10 
Y=X*4+3.45 
Y=rnorm(100)*0.29*Y+Y
```
## a) Plot Y against X. Include a screenshot of the plot in your submission. Using the File menu you can save the graph as a picture on your computer.  Based on the plot do you think we can fit a linear model to explain Y based on X? (5 Marks) 
```{r}
plot(X,Y, xlab = "X (units)", ylab = "Y (units)", main = "Plot of Y vs X for Question 1.a", col = "red")
```
From the graph shown above you can see a positive linear trend between X and Y,
 
## b) Construct a simple linear model of Y based on X. Write the equation that explains Y based on X. What is the accuracy of this model? (5 Marks) 
```{r}
Model <- lm(Y~X)
Model$coefficients
```
The formula to explain Y based on X from our linear model is: Y = 3.6108*X + 4.4655.

## c) How  the  Coefficient  of  Determination,  R2,  of  the  model  above  is  related  to  the  correlation coefficient of X and Y? (5 marks) 
```{r}
summary(Model)
```
The R2 is 0.6517, that means 65% of the variability of Y is captured by X.


### 2) We  will use the ‘mtcars’ dataset for this question. The dataset is already included in your R distribution. The dataset shows some of the characteristics of different cars. The following shows few samples (i.e. the first 6 rows) of the dataset.
```{r}
# shows first 6 rows.
head(mtcars)
```

## a) James wants to buy a car.  He and his friend, Chris, have different opinions about the Horse Power  (hp)  of  cars. James think the  weight  of  a  car  (wt)  can  be  used  to  estimate  the  Horse Power of the car while Chris thinks the fuel consumption expressed in Mile Per Gallon (mpg), is a better estimator of the (hp).  Who do you think is right? Construct simple linear models using mtcars data to answer the question. (10 marks) 
```{r}
# Creates a linear model for weight vs horsepower and displays a plot of the points
Model2 = lm(hp~wt, data = mtcars)
summary(Model2)
plot(mtcars$wt,mtcars$hp, xlab = "Weight (wt)", ylab = "Horsepower (hp)", main = "Plot of Weight vs Horsepower for Question 2.a", col = "blue")
```
From this linear model we can see that weight results in a model that accounts for 43.39% of the variation in horsepower.

```{r}
# Creates a linear model for mpg vs horsepower and displays a plot of the points
Model3 = lm(hp~mpg, data = mtcars)
summary(Model3)
plot(mtcars$mpg,mtcars$hp, xlab = "Miles per Gallon (mpg)", ylab = "Horsepower (hp)", main = "Plot of Miles Per Gallon vs Horsepower for Question 2.a", col = "blue")
```
From this linear model we can see that fuel efficiency results in a model that accounts for 60.24% of the variation in horsepower. Therefore, fuel efficiency (mpg) is considered statistically significant in this model.

## b) Build a model that uses the number of cylinders (cyl) and the mile per gallon (mpg) values of a car to predict the car Horse Power (hp). Using this model, what is the estimated Horse Power of a car with 4 calendar and mpg of 22? (10 mark) 
```{r}
# Shows which variables are factor or numeric
str(mtcars)

# Convert cylinder into a factor.
mtcars$cyl = as.factor(mtcars$cyl)
str(mtcars)
```
# A multiple regression model is utilized in this case to build a model that represents horsepower as a result of cylinders and miles per gallon.
```{r}
Model5 = lm(hp~cyl+mpg, data = mtcars)
summary(Model5)

# Predict the estimated horse power of a car with 4 cylinders and 22 mpg
predict(Model5, data.frame(mpg = c(22), cyl = c("4")))
```
The estimated Horse Power of a car with 4 calendar and mpg of 22 is 98.15%


### 3) For this question, we are going to use BostonHousing dataset. The dataset is in ‘mlbench’ package, so we first need to instal the package, call the library and the load the dataset using the following commands
```{r}
library(mlbench)
data(BostonHousing)
```

## a) Build a model to estimate the median value of owner-occupied homes (medv)based on the following  variables:  crime  crate  (crim),  proportion  of  residential  land  zoned  for  lots  over 25,000  sq.ft  (zn),  the  local  pupil-teacher  ratio  (ptratio)  and  weather  the  whether  the  tract bounds Chas River(chas). Is this an accurate model? (Hint check R2 )
```{r}
# Check the dataset to see which variable are considered numeric and which are factors.
str(BostonHousing)
```

```{r}
# Create a linear model for median value based on crim, zn, ptratio, and chas.
Model6 = lm(medv~crim+zn+ptratio+chas, data = BostonHousing)
summary(Model6)
```
The R2 value in this model (crim, zn, ptratio, and chas) captured is 35.99% of the variability in median home value. This is a weak model in terms of accuracy and can be improved by adding more variables into the model.

## b) Use the estimated coefficient to answer these questions?   
 I) Imagine  two  houses that  are  identical  in  all aspects  but  one  bounds  the  Chas  River and the other does not. Which one is more expensive and by how much?

Based on the coefficients, the resulting formula from our model is:

medv = 49.91868 - 0.26018*crim + 0.07073*zn - 1.49367*ptratio + 4.58393*chas1

Therefore, if the only difference between two houses is that one borders the Chas River, then we would only focus on the chas variable coefficient. The house that borders the river would be $4,583.93 more than the one that does not.

4.58393 (coeff of chas) * 1 (value of chas) * 1000 (medv in $1,000 units) = $4,583.93

II)	Imagine two houses that are identical in all aspects but in the neighborhood of one of them the pupil-teacher ratio is 15 and in the other one is 18. Which one is more expensive and by how much? (Golden Question: 10 extra marks if you answer)

Based on the coefficients, the resulting formula from our model is:

medv = 49.91868 - 0.26018*crim + 0.07073*zn - 1.49367*ptratio + 4.58393*chas1

Therefore, if the only difference between two houses is the pupil-teacher ratio, then we would only focus on the ptratio variable coefficient. As a result, the house with the smaller pupil-teacher ratio value would be more expensive, because the coefficient is found to be negative in our model. The difference in values between the houses would be:

-1.49367 (coeff of ptratio) * 0.03 (difference between ptratio values) * 1000 (medv in $1,000 units) = $44.81

Therefore, the house with the lower pupil-teacher ratio would be $44.81 more expensive based on our model.

# c)	Which of the variables are statistically important (i.e. related to the house price)? Hint: use the p-values of the coefficients to answer.

Based on the model constructed from these variables, all of the variables (crim, zn, ptratio, and chas) were found to be statistically significant. This is true because all of the p-values calculated from our model at below the 0.05 threshold value for significance.

# d)	Use the anova analysis and determine the order of importance of these four variables. 

```{r}
# Returns the ANOVA results for the model used in this problem
anova(Model6)
```

Based on the ANOVA values returned, the order of importance of these variables are:

1. "crim" - accounts for 15.08% of variability in the model
2. "ptratio" - accounts for 11.02% of variability in the model
3. "zn" - accounts for 8.32% of variability in the model
4. "chas" - accounts for 1.56% of variability in the model

Additionally, the residuals in this model still account for 64.01% of variability in the model, so there is still a lot of room for improvement in the accuracy of this model.
 







